---
title: "Traktplott for variabler fra NKR-Degenerativ nakke"
author: "Lena Ringstad Olsen, basert på kode fra Henrik Hedevik og Bjørnar Berg"
date: "`r Sys.Date()`"
output: html_document
note: Requires prepared NORspine data restricted to time-period and surgical cases of interest (long format)
variables needed: success (dep.var), clinic (shnavn), covariates: alder, kjonn, asa, prevsurg(binar), oswtotpre
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Installation of packages

```{r packages, echo=FALSE}
#if(!require(foreign)) { install.packages("foreign"); require(foreign)}
#library(foreign)
if(!require(ggplot2)) { install.packages("ggplot2"); require(ggplot2)}
#library(ggplot2)
if(!require(dplyr)) { install.packages("dplyr"); require(dplyr)}
#library(dplyr)
if(!require(broom)) { install.packages("broom"); require(broom)}
#library(broom)
if(!require(data.table)) { install.packages("data.table"); require(data.table)}
#library(data.table)
if(!require(emmeans)) { install.packages("emmeans"); require(emmeans)}
# library(emmeans)
# library(haven) 
# library(ggrepel)



# Traktplott for "suksess" (%)
#' Generell traktplottfunksjon. Krever tilpasset datasett
#'
#' @param data dataramme, ryggdata med tilrettelagte variabler. Må inneholde
#' xaksetxt tekst på x-aksen. 
#' yaksetxt tekst på x-aksen. 
#' tittel tittel
#'
#' @return
#' @export
#'
Traktplott <- function(data_aggregert = LPHD_Clinic,
                       xaksetxt = 'Antall operasjoner',
                       yaksetxt = 'Suksess',
                       tittel = 'Angi tittel') {
  x_maks <- ceiling(max(data_aggregert$xvar)/100)*100
  library(ggplot2)
  ggplot2::ggplot(data = data_aggregert, aes(x = xvar, group = OverallRate)) +
     geom_smooth(aes(y=konf_opp), se=FALSE, linetype="solid", color="red", linewidth=0.8) +
     geom_smooth(aes(y=konf_ned), se=FALSE, linetype="solid", color="red", linewidth=0.8) +
     geom_smooth(aes(y=OverallRate), se=FALSE, color="darkgray", linewidth=0.8) +
     geom_point(aes(y = Rate), color = "black", size = 2) +
     ggrepel::geom_text_repel(aes(y = Rate, label = SykehusNavn), 
                #              label.size = 
                color = "black", 
                box.padding = 0.4, # Adds padding between labels and points
                point.padding = 0.2, # Adds padding between points and lines
                max.overlaps = 10, # Adjusts the number of overlapping labels allowed
                min.segment.length = 0.2, # Avoids long connecting lines
                segment.color = "gray") + # Sets line color
  theme_classic() +
  scale_x_continuous(breaks = seq(0, x_maks, by = 100), limits = c(0, x_maks)) +
  scale_y_continuous(breaks=seq(0.20, 1, by=0.05), labels=scales::percent) +
  labs(x =   'Antall operasjoner',
       y = 'Suksess',
       title = tittel)
}

```
### Loading data LR

David-modell:
Alder >60	Ja/nei
Fedme (BMI)	Ja/nei
Ikke i arbeid (dikotomisert)	Ja/nei
Tungt fysisk arbeid	Ja/nei
Lavt utdanningsnivå	Ja/nei
SykdomDiabetes	Ja/nei
Ikke norsk morsmål	Ja/nei
Røyker	Ja/nei
Tidligere nakkeoperert	Ja/nei
Operert mer enn 1 nivå	Ja/nei
Søkt erstatning	Ja/nei
Varighet av armsmerter  3-12 mnd	Ja/nei
Varighet av armsmerter  >12 mnd	Ja/nei
NDI hodepine	6 kategorier

Full modell:
Alder >60	Ja/nei
Kjønn	 
Fedme (BMI)	Ja/nei
Ikke i arbeid (dikotomisert)	Ja/nei
Tungt fysisk arbeid	Ja/nei
Lavt utdanningsnivå	Ja/nei
Sykdom Diabetes	Ja/nei
ASA> II	Ja/nei
Ikke norsk morsmål	Ja/nei
Røyker	Ja/nei
Tidligere nakkeoperert	Ja/nei
Operert mer enn 1 nivå	Ja/nei
Søkt erstatning	Ja/nei
Varighet av armsmerter  3-12 mnd	Ja/nei
Varighet av armsmerter  >12 mnd	Ja/nei
Nakkesmerter> Armsmerter (NRS)	Ja/nei
Angst depresjon (EQ-5D)	Nei mot resten
NDI levels (<40, 40-60, >60),	 
NRS arm levels (0-5, 6-7, >8),	 
NDI hodepine	6 kategorier

For each institution, the percentage of negative outcomes based on each of the three models was calculated (mean of the outcome). 
Consequently, the predicted probability for each unit based on the given case mix of the patient cohort, 
was subtracted from the mean of the cutoff variable. The resulting value represents 
the adjusted proportion of negative outcome. Positive adjusted values indicate a resulting 
performance worse than the model predicted and negative (minus) adjusted values a performance better than the model predicted.

Failure:
For adjustment the modelled probability for the outcome was subtracted from the achieved outcome. 
Negative values indicated performance better than model prediction.

```{r data, echo=FALSE}

source("C:/Users/lro2402unn/RegistreGIT/nakke/dev/sysSetenv.R")
library(nakke)
NakkeDataRaa <- NakkeHentRegData(datoFra = '2014-01-01',datoTil = '2023-12-31') #13927
RegData <- NakkePreprosess(NakkeDataRaa)

  RegData <- NakkeUtvalgEnh(RegData = RegData, myelopati = 0)$RegData
  RegData <- RegData[RegData$Inngrep %in% c(1,3), ]
  RegData <- NakkeVarTilrettelegg(RegData = RegData, valgtVar = 'NDIendr12mnd35pst')$RegData

  
#   variabler <- c('SykehusNavn', 'Variabel', 'OprDato')
#   variablerTilAn <- c('Alder', 'ASAgrad', 'TidlOpr', 'Roker', 'Inngrep', 'Utdanning', 'ErMann', 'NDIscorePreOp',
#                      'BMI', 'ArbeidstausPreOp', 'SymptVarighetArmer', 'EqAngstPreOp', 'AntallNivaaOpr')
# RegData <- RegData[ ,c(variabler, variablerTilAn)]
  
# Kovariater («independent variables»):
# Alder, kjønn (ErMann), preoperativ skår (NDIscorePreOp)
RegData$Alder60 <- ifelse(RegData$Alder > 60, 1,0)
RegData$ASAover2 <- as.numeric(RegData$ASAgrad>2)  # ASAgrad: >2 (ja/nei) dikotom variabel
RegData$ErTidlOpr <- as.numeric(RegData$TidlOpr %in% 1:3) # TidlOpr (ja/nei)
RegData <- RegData[which(RegData$Roker %in% 0:1), ] # Roker (1ja, 0 nei), 9=NULL
RegData <- RegData[which(RegData$Utdanning %in% 1:5), ] # Utdanning (1-3=1 (lav utdanning) og 4-5= 2 (Høy utdanning))
RegData$UtdHoy <- ifelse(RegData$Utdanning %in% 4:5, 1,0)
RegData$Fedme <- ifelse(RegData$BMI >= 30, 1,0)  # BMI ≥ 30.0=1 (fedme) og <30= 0 (ikke fedme)
RegData <- RegData[which(RegData$ArbeidstausPreOp %in% 1:10), ]
RegData$Sykepenger <- ifelse(RegData$ArbeidstausPreOp %in% 6:10, 1,0) # ArbeidstausPreOp (6-10=1 (mottar sykepenger) og 1-5=0 (mottar ikke sykepenger))
RegData <- RegData[which(RegData$SymptVarighetArmer %in% 2:5), ] #SymptVarighetArmer (1=NULL, 2-3=1 (varighet under ett år) og 4-5= 2 (symtomvarighet over ett år))
RegData$ArmsmMin3mnd <- ifelse(RegData$SymptVarighetArmer %in% 3:5, 1,0)
RegData$ArmsmMin12mnd <- ifelse(RegData$SymptVarighetArmer %in% 4:5, 1,0)
RegData <- RegData[which(RegData$AntallNivaaOpr > 0), ]  # AntallNivaaOpr (0=NULL, 1=1 (ett nivå) og >1=2 (mer enn ett nivå))
RegData$OpFlereNivaa <- ifelse(RegData$AntallNivaaOpr >1, 1,0)
RegData <- RegData[RegData$YrkeFysiskeKrav %in% 1:5, ]
RegData$TungtArb <- ifelse(RegData$YrkeFysiskeKrav==3, 1, 0)
RegData$Norsk <- ifelse(RegData$Morsmal==1, 1, 0)

RegData <- RegData |>
  dplyr::mutate(
    NDI3kat = dplyr::case_when(
      NDIscorePreOp < 40 ~ 1, # '<40',
      NDIscorePreOp >= 40 & NDIscorePreOp <= 60 ~ 2, # '40-60',
      NDIscorePreOp > 60 ~ 3), # '>60'),
    ArmNRSkat = dplyr::case_when(
      NRSsmerteArm12mnd %in% 0:5 ~ 1, # '0-5', 
      NRSsmerteArm12mnd %in% 6:7 ~ 2, #'6-7', 
      NRSsmerteArm12mnd >8  ~ 3), # '>8'),	 
    NDIhodepinePreOp = dplyr::na_if(NDIhodepinePreOp, 9),
    EqAngstPreOp = dplyr::na_if(EqAngstPreOp, 9),
    Angst = ifelse(EqAngstPreOp >1, 1,0),
    ErstatningPreOp = dplyr::na_if(RegData$ErstatningPreOp, 9),
    SoktErstatn = dplyr::case_when(ErstatningPreOp %in% c(1,3) ~ 1,
                            ErstatningPreOp %in% c(2,4) ~ 0), #Kode 1:4,9: 'Ja', 'Nei', 'Planlegger', 'Innvilget', 'Ukjent'
    MerSmerteNakke = NRSsmerteNakke12mnd > NRSsmerteArm12mnd
)
RegData <- RegData |> tidyr::drop_na(NDIhodepinePreOp)
 

# Slå sammen private:
  indPriv <- which(RegData$ReshId %in% c(999998, 999920, 110771, 107511, 4211880, 4211879))
  RegData$SykehusNavn[indPriv] <- 'Private'
 #Oslofjordklinikken, Oslofjordklinikken Vest, Volvat, Aleris Oslo, Aleris Nesttun, Aleris Stavanger

RegDataStart <- RegData

```

### Beregner data

```{r LPHD_LogReg, echo=FALSE}

# RegData <- RegData |>
#   mutate(across(
#       c(Fedme, Sykepenger, TungtArb, UtdHoy, SykDiabetesMellitus , Roker,
#         Norsk, ErTidlOpr, OpFlereNivaa, SoktErstatn, ArmsmMin3mnd, ArmsmMin12mnd, 
#         MerSmerteNakke, Angst, NDI3kat, ArmNRSkat),
#       ~ forcats::fct_na_value_to_level(as.factor(.x), level = NA)))


# Calculate Clinic Volume and Satisfied patients
RegData <- RegData |> dplyr::group_by(SykehusNavn) |> 
  dplyr::mutate(ClinicVolume = dplyr::n(),
         AntSuccessPerSh := sum(Variabel)) |>
  dplyr::ungroup()

# Filtrerer bort enheter med få registreringer
LPHD <- RegData[RegData$ClinicVolume > 20, ]

# Logistic Regression model on Success
#LPHD_Model <- glm(success ~ clinic + alder + kjonn + prevsurg + asa + oswtotpre, family = binomial, data = LPHD)

#Ingen justering
tittel <- 'Ingen justering'
LPHD_Model <- glm(Variabel ~ SykehusNavn, family = binomial, data = LPHD)
summary(LPHD_Model)
data_aggregert <- tilretteleggForTrakt(LPHD_Model)
trakt <- Traktplott(data_aggregert = data_aggregert,
           tittel = tittel)
ggsave(paste0('Traktplott, ',tittel,".pdf"), width = 10, height = 6) #plot = PlotRateAdj, 

#Alder og kjønn
tittel <- 'Justert for alder og kjønn (ikke signifikante)'
LPHD_Model <- glm(Variabel ~ SykehusNavn + Alder + ErMann, family = binomial, data = LPHD)
summary(LPHD_Model)
data_aggregert <- tilretteleggForTrakt(LPHD_Model)
Traktplott(data_aggregert = data_aggregert,
           tittel = tittel)
ggsave(paste0('Traktplott, ',tittel,".pdf"), width = 10, height = 6) #plot = PlotRateAdj, 

#Tores modell:
tittel <- 'Justert, Tores modell (Sign var, 5pst)'
LPHD_Model <- glm(Variabel ~ SykehusNavn + ErTidlOpr + Roker + 
             Sykepenger + ArmsmMin12mnd + Angst + OpFlereNivaa
             #Fedme + ASAover2 + Alder + NDIscorePreOp +  ErMann + UtdHoy + 
             , family = binomial(link = "logit"), data = LPHD)
summary(LPHD_Model)
data_aggregert <- tilretteleggForTrakt(LPHD_Model)
Traktplott(data_aggregert = data_aggregert,
           tittel = tittel)
ggsave(paste0('Traktplott, ',tittel,".pdf"), width = 10, height = 6) #plot = PlotRateAdj, 
                  

#Davids modell:
tittel <- 'Justert, Davids modell (Sign var, 5pst)'
LPHD_Model <- glm(Variabel ~ SykehusNavn + Sykepenger  + #Fedme + Alder60 + 
                      TungtArb + UtdHoy + SykDiabetesMellitus + Roker +
                      Norsk + ErTidlOpr + OpFlereNivaa + SoktErstatn + 
                      ArmsmMin3mnd + ArmsmMin12mnd + NDIhodepinePreOp
                    , family = binomial, data = LPHD )
#                  , na.action = 'na.exclude')
summary(LPHD_Model)
data_aggregert <- tilretteleggForTrakt(LPHD_Model)
Traktplott(data_aggregert = data_aggregert,
           tittel = tittel)
ggsave(paste0('Traktplott, ',tittel,".pdf"), width = 10, height = 6) #plot = PlotRateAdj, 

# Unadjusted, the range of non-successful outcomes between hospitals was up to 25%. After risk adjustment, 
# inter-hospital differences diminished markedly (range 1.9%–3.6%). Only private sector status remained 
# statistically significant in all three models. High odds ratios (>2.0) for non-success were observed for
# medical litigation, prior cervical surgery, duration of arm pain >12 months, and headache as a dominant symptom.

#Full modell:
tittel <- 'Justert, full Davidmodell (sign. var 5pst)'
  LPHD_Model <- glm(Variabel ~ SykehusNavn + 
                      Roker  + ErTidlOpr + OpFlereNivaa + 
                      ArmsmMin12mnd  + MerSmerteNakke + NDI3kat + ArmNRSkat + NDIhodepinePreOp 
                      #ErMann +TungtArb + UtdHoy + Fedme +  ArmsmMin3mnd + SoktErstatn + 
                      #Alder60 + Norsk + SykDiabetesMellitus + Sykepenger+  Angst
                    , family = binomial(link = "logit"), data = LPHD)
                    #, na.action = 'na.exclude')
  
summary(LPHD_Model) #Alle signifikante på 5%-nivå (NDIkat, bare >60 sign)

data_aggregert <- tilretteleggForTrakt(LPHD_Model)
Traktplott(data_aggregert = data_aggregert,
           tittel = tittel)
ggsave(paste0('Traktplott, ',tittel,".pdf"), width = 10, height = 6) #plot = PlotRateAdj, 



#--------Predikert verdi:-----------
# LPHD$predikert <- predict(object=LPHD_Model, type = "response") #, na.action = na.exclude)
# 
# hospital_summary <- LPHD %>%
#   group_by(SykehusNavn) %>%
#   summarise(
#     Observed = mean(Variabel, na.rm = T),
#     Expected = mean(predikert, na.rm = T),
#     N = n(),
#     OE_ratio = Observed / Expected
#   )
# LPHD <- merge(LPHD, LPHD_Model$fitted.values, by = "SykehusNavn", all.x = TRUE)
# head(LPHD_Model$fitted.values)




#' Beregne EMM og tilrettelegge for traktplott
#'
#' @param LPHD_Model 
#'
#' @return
tilretteleggForTrakt <- function(LPHD_Model){
# Calculate estimated marginal means for Clinic on the response scale (EMMs are also known as least-squares means)
LPHD_EMMs <- emmeans::emmeans(LPHD_Model, ~ SykehusNavn, type = "response", rg.limit=30000)

# View the EMMs on the response scale
summary(LPHD_EMMs)

# Merge EMMs with the original data table
LPHD_Prob <- data.table::as.data.table(summary(LPHD_EMMs))
LPHD <- merge(LPHD, LPHD_Prob, by = "SykehusNavn", all.x = TRUE)
LPHD$AntSuccessPerShAdj <- LPHD$ClinicVolume*LPHD$prob

# Exclude duplicate cases in Clinic 
# LPHD_Clinic <- LPHD[, .(clinic, ClinicVolume, AntSuccessPerSh, AntSuccessPerShAdj)][!duplicated(clinic)]
LPHD_Clinic <- LPHD[!duplicated(LPHD$SykehusNavn), c('SykehusNavn', 'ClinicVolume', 'AntSuccessPerSh', 'AntSuccessPerShAdj')]

# Calculate Satisfaction Rates
LPHD_Clinic <- LPHD_Clinic |>
  dplyr::mutate(Rate = AntSuccessPerSh/ClinicVolume, #Andel suksess observert
                OverallRate = sum(AntSuccessPerSh)/sum(ClinicVolume),
                se = sqrt(OverallRate * ((1 - OverallRate)/ClinicVolume)),
                lcl95 =  OverallRate - (1.96 * se),
                ucl95 =  OverallRate + (1.96 * se),
                RateAdj = AntSuccessPerShAdj/ClinicVolume, #Andel suksess estimert
                OverallRateAdj = sum(AntSuccessPerShAdj)/sum(ClinicVolume), 
                seAdj = sqrt(OverallRateAdj * ((1 - OverallRateAdj)/ClinicVolume)),
                lcl95Adj =  OverallRateAdj - (1.96 * seAdj),
                ucl95Adj =  OverallRateAdj + (1.96 * seAdj))
#View(LPHD_Clinic)
#LPHD_Clinic_Full <- LPHD_Clinic
#Standardiserer inngangsvariabler til plottet
data_aggregert <- LPHD_Clinic |> 
  dplyr::mutate(xvar = ClinicVolume, 
                OverallRate = OverallRateAdj,
                Rate = RateAdj,
                konf_opp = lcl95Adj, 
                konf_ned = ucl95Adj)
return(data_aggregert)
}
```

### Funnel plot Satisfaction (%) LPHD

```{r LPHD_Plot, echo=FALSE}




```

